---
title: "060822_isomiRMap"
output: html_document
date: '2022-06-09'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/RNAseq/exc_exp_text/")
library(plyr)
library(DESeq2)
library(stringi)
library(purrr)
library(dplyr)
library(tidyverse)
library(rlist)

control <- "control"
stroke <- "stroke"
de <- data.frame(row.names = c("1","3","4","8","9","13","14","18","19","20","21","22","23","24","30","33","34","35","36","37","38","41","43","44","47","48","50","53","54","55","56","57","61","62","63","65","69","70","72","73","74","76","78","79","81","84","86","87","88","91","92","93","94","95","96","97","98","107","110","113","115","117","118","121","122","124","126","131","133","134","135","137","140","141","142"), condition = c(control,control,control,stroke,control,control,stroke,control,stroke,control,control,control,control,control,stroke,stroke,stroke,stroke,stroke,stroke,stroke,stroke,control,stroke,control,control,stroke,control,control,stroke,control,control,control,stroke, control,control,control,stroke,stroke,stroke,control,stroke,control,stroke,stroke,control,stroke,control,control,control,control,stroke,stroke,stroke,stroke,stroke,control,control,control,control,stroke,stroke,control,control,control,stroke,control,stroke,stroke,control,control,stroke,control,control,control), stringsAsFactors = TRUE)




```{r}

```


```{r}
## generate list of samples + conditions (ordered by sample) 

map_list <- list.files(path = ".", pattern="txt", full.names = TRUE)
row_name <- row.names(de)

new_names<- list()
for (i in 1:length(map_list)){
  a_name <- paste0(de$condition[i], row_name[i])
  new_names[[i]] <- a_name
}
new_names <- as_vector(new_names)

col_conditions <- data.frame(condition = de, row.names = new_names)

## pull sample master table and merge to dataframe with license plate, counts, and corresponding condition 

map_dfs <- list()
for (i in 1:length(map_list)){
  temp_name <- paste0(de$condition[i], row_name[i])
  map_v <- read.delim(map_list[i], header = TRUE, skip = 6)
  df_v <- data.frame(names = map_v$License.Plate, temp_name = map_v$Unnormalized.read.counts)
  names(df_v)[2] <- temp_name
  map_dfs[[i]] <- df_v
}

## merge all count dataframes by license plate and tidy data

all_of_them <- Reduce(function(x, y) merge(x, y, all=TRUE), map_dfs)
row.names(all_of_them) <- all_of_them$names
all_of_them$names <- NULL
all_of_them[is.na(all_of_them)] <- 0

## create DEseq data object 

big_ids <- DESeqDataSetFromMatrix(all_of_them, colData = col_conditions, design = ~condition)

```

```{r}
## Run DEseq on full count table (with license plates)

big_new_ids <- DESeq(big_ids)

keep <- rowSums(counts(big_new_ids)) >= 10
big_new_ids <- big_new_ids[keep,]

big_res <- results(big_new_ids, contrast = c("condition", "stroke", "control"), lfcThreshold = 0.2)
p.threshold <- 0.05

big_res$threshold <- as.logical(big_res$padj < p.threshold)
genes.deseq_big <- row.names(big_res)[which(big_res$threshold)]


all_genes_big_res <- data.frame(big_res$padj, big_res$log2FoldChange, temp_name = row.names(big_res))
p_sort_big_res <- all_genes_big_res[order(all_genes_big_res$big_res.padj),]

plotCounts( big_new_ids, gene="iso-20-HBS2DF4I", intgroup="condition")

resLFC_apeglm_mat <- lfcShrink(big_new_ids, coef="condition_stroke_vs_control", type="apeglm")
plotMA(resLFC_apeglm_mat, xlim=c(-1e2,5000),ylim=c(-1,10000))

write.table(genes.deseq_big, "lfc0.2_p0.05_DE_isomirs.txt", row.names = FALSE, col.names = FALSE, quote = FALSE)
```

```{r}
## BROKEN

## gather all counts by miR ID

eff <- list()
for (i in 1:length(map_list)){
  temp_name <- paste0(de$condition[i], row_name[i])
  map_v <- read.delim(map_list[i], header = TRUE, skip = 6)
  df_v <- data.frame(names = map_v$Mature.meta.data..bracket.delimited.per.hairpin., temp_name = map_v$Unnormalized.read.counts)
  hope <- vector()
  for (j in 1:nrow(df_v)){
    x <- df_v$names[j]
    peace <- stri_extract(x, regex="(?<=\\[)(.*?)(?=,)")
    hope[j] <- peace
  }
  df_v$names<- hope
  df_v <- na.omit(df_v)
  new_v <- data.frame(names = df_v$names, temp_name = df_v$temp_name)
  names(df_v)[2] <- temp_name
  eff[[i]] <- new_v
}

the_names <- Reduce(function(x, y) merge(x, y, all=TRUE), eff)

```

```{r}
## gather license plate IDs + miR name IDs

eff <- list()
for (i in 1:length(map_list)){
  temp_name <- paste0(de$condition[i], row_name[i])
  map_v <- read.delim(map_list[i], header = TRUE, skip = 6)
  df_v <- data.frame(names = map_v$Mature.meta.data..bracket.delimited.per.hairpin., temp_name = map_v$License.Plate)
  eff[[i]] <- df_v
}

the_names <- Reduce(function(x, y) merge(x, y, all=TRUE), eff)

## merge miR name IDs to cooresponding license plate IDs of DE miRs 

names(p_sort_big_res)[2] <- "temp_name"
all_names <- inner_join(p_sort_big_res, the_names, by="temp_name")

```

```{r}
## pull all unique mature miRs

hope <- vector()

for (j in 1:nrow(all_names)){
  x <- all_names$names[j]
  if(stri_detect(x, regex = "MIMA")){
    peace <- stri_extract(x, regex="(?<=\\&)(.*?)(?=\\&)")
    hope[[j]] <- peace
  }
  else if(stri_detect(x, regex= "\\[m-")){
    peace <- stri_extract(x, regex="(?<=\\m-)(.*?)(?=\\|)")
    hope[[j]] <- peace
  }
  else{
    peace <- stri_extract(x, regex="(?<=\\MD2.)(.*?)(?=\\-miR)")
    hope[[j]] <- peace
  }
}

hope <- tibble(hope) %>% 
  unique() %>% 
  drop_na()
#10053&19
#hsa-miR-144-3p
#hsa-miR-142-5p

generate_iso_data <- function(isomiR_name){
  name_group <- the_names[grep(isomiR_name, the_names$names),]
  pull_license_plates <- all_of_them[grep(name_group[3,2], row.names(all_of_them)),]

  count_group <- list()
  for (i in 1:nrow(name_group)){
   pull_license_plates <- all_of_them[grep(name_group[i,2], row.names(all_of_them)),]
   count_group[[i]] <- pull_license_plates
  }

  bound_group <- list.rbind(count_group)

  bound_group$temp_name <- row.names(bound_group)
  bound_group <- inner_join(bound_group, the_names, by="temp_name")
  
  isomiR_mods <- list()

  for (i in 1:nrow(bound_group)){
     x <- bound_group$names[i]
     temp_df <- data.frame(five = NA, three = NA, add = NA)
     five_end <- stri_extract(x, regex="(?<=s\\|)([+-]?\\d{1,2})")
     temp_df$five <- five_end
     three_end <- stri_extract(x, regex="(?<=s\\|[+-]?\\d{0,1}\\|)([+-]?\\d{1,2})")
     temp_df$three <- three_end
     addition <- stri_extract(x, regex="(?<=\\()(.*?)(?=\\))")
     temp_df$add <- addition
     temp_df[is.na(temp_df)] <- 0
     isomiR_mods[[i]] <- temp_df
  }

  all_mods <- list.rbind(isomiR_mods)
  all_mods$temp_name <- bound_group$temp_name

  bound_group$row_sums <- rowSums(bound_group[,1:75])
  bound_group$control_sums <- rowSums(bound_group[, grepl("control", names(bound_group))])
  bound_group$stroke_sums <- rowSums(bound_group[, grepl("stroke", names(bound_group))])
  bound_group <- inner_join(bound_group, all_mods, by="temp_name")

  shift_test <- function(dataframe, number){
    if((dataframe$five[number] == dataframe$three[number]) &
      (dataframe$five[number] != 0))
     {return(TRUE)}
    else (return(FALSE))
  }
  start_site_iso <- function(dataframe, number){
   if(dataframe$five[number] != 0){
     return(TRUE)
    } else (return(FALSE))
  }
  end_site_iso <- function(dataframe, number){
    if(dataframe$three[number] != 0){
     return(TRUE)
   } else (return(FALSE))
  }
  nta_iso <- function(dataframe, number){
    if(dataframe$add[number] != 0){
      return(TRUE)
    } else (return(FALSE))
  }

  five_site <- list()
  three_site <- list()
  shifts <- list()
  ntas <- list()
  for(i in 1:length(bound_group$five)){
    each_shift <- shift_test(bound_group, i)
    each_five <- start_site_iso(bound_group, i)
    each_three <- end_site_iso(bound_group,i)
    each_add <- nta_iso(bound_group,i)
    shifts[[i]] <- each_shift 
    five_site[[i]] <- each_five
    three_site[[i]] <- each_three
    ntas[[i]] <- each_add
  }

  bound_group$shift <- shifts
  bound_group$start_site <- five_site
  bound_group$end_site <- three_site
  bound_group$add_isos <- ntas

  bound_group <- left_join(bound_group, all_genes_big_res, by="temp_name")
  
  return(bound_group)

}

hsa_miR_144_3p_data <- generate_iso_data("hsa-miR-144-3p")
hsa_miR_142_5p_data <- generate_iso_data("hsa-miR-142-5p")
hsa_miR_140_5p_data <- generate_iso_data("hsa-miR-140-5p")

#write.table(bound_group$temp_name, "m-10053&19_isomirs.txt", row.names = FALSE, col.names = FALSE, quote = FALSE)

```


```{r}
## test chunk

#df_all <- join_all(map_dfs, by = "names")
#test1111 <- read.delim(map_list[1], header = TRUE, skip = 6)
#test1112 <- read.delim(map_list[2], header = TRUE, skip = 6)

#df1 <- data.frame(names = test1111$License.Plate, test1111$Unnormalized.read.counts)
#df2 <- data.frame(names = test1112$License.Plate, test1112$Unnormalized.read.counts)
#df1_2 <- join(df2, df1, by = "names")

for (i in 1:length(map_list)){
  map_v <- read.delim(map_list[i], header = TRUE, skip = 6) 
  df_v <- data.frame(names = map_v$Mature.meta.data..bracket.delimited.per.hairpin., temp_name = map_v$Unnormalized.read.counts)
}

hope <- vector()
for (i in 1:nrow(map_v)){
  x <- map_v$names[i]
  peace <- stri_extract(x, regex="(?<=\\[)(.*?)(?=,)")
  hope[i] <- peace
}
map_v$names<- hope

x <- map_v$Mature.meta.data..bracket.delimited.per.hairpin.[2]
peace <- stri_extract(x, regex="(?<=\\[)(.*?)(?=,)")
```

